# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
We assume that the data is in the same folder. 
We load data, and just convert columns to correct types

```{r}
file<-read.csv("activity.csv", stringsAsFactors=FALSE)
file$steps = as.numeric(file$steps)
file$date=as.Date(file$date)
file$interval=as.numeric(file$interval)
```


## What is mean total number of steps taken per day?
We group data by Date, and calculate Sum of steps over specific dates.

```{r}
stepsPerDay <- tapply(file$steps,file$date,sum)
summary(stepsPerDay)
```
Let's calculate mean and median over these steps
```{r}
meanStepsPerDay <- mean(na.omit(as.vector(stepsPerDay)))
medianStepsPerDay <- median(na.omit(as.vector(stepsPerDay)))
```
Mean steps per day: `r meanStepsPerDay`, median: `r medianStepsPerDay`.

## What is the average daily activity pattern?

Here is data for average daily activity pattern, as average steps by interval
```{r}
patternStepsPerInterval = tapply(file$steps,file$interval,mean, na.rm=TRUE)
summary(patternStepsPerInterval)
patternStepsNames <-rownames(patternStepsPerInterval)
plot(patternStepsNames, patternStepsPerInterval,type="l",
   xlab="Interval #",ylab="Avg steps across days")

```
Calculating maximum steps per day
```{r}
maxpatternStepsPerInterval = max(patternStepsPerInterval) 
maxpatternStepsAtInterval = patternStepsNames[which.max(patternStepsPerInterval)]
```
Max pattern steps per day on average: `r maxpatternStepsPerInterval` occur at interval `r maxpatternStepsAtInterval`.
## Inputing missing values
Calculating missing values
```{r}
message(paste("This dataset has", 
  nrow(file[is.na( file$steps),]), 
  "missing values"))
```
We fill missing values with mean for that 5-minute interval, calculated above

```{r}
fullData =data.frame(steps= file$steps,date=file$date,interval=file$interval)
fullData$steps <- with(fullData,ifelse(is.na(steps),
  patternStepsPerInterval[as.character(interval)],
  steps))
```

Now, here's usage pattern for total steps per day, taking into account missing values
```{r}
fullStepsPerDay = tapply(fullData$steps,fullData$date,sum)
head(fullStepsPerDay)
hist(fullStepsPerDay,xlab="Total steps (with replaced values)",main="Histogram (with replaced missed values)")
```

This method of filling missed values does not change mean of steps per day compared with original data. Median, however, will be changed
```{r}
fullmeanStepsPerDay =mean(fullStepsPerDay)
fullmedianStepsPerDay = median(fullStepsPerDay)
```
After filling data, mean is different from original dataset by `r fullmeanStepsPerDay - meanStepsPerDay`, median is different by `r fullmedianStepsPerDay-medianStepsPerDay`.



## Are there differences in activity patterns between weekdays and weekends?
We fill dataset with information on weekdays, and aggregate it over interval # and (isWeekend) function
```{r}
isWeekend <- function(d) {
  val <- weekdays(as.Date(d), abbreviate=TRUE)
  if(val=='Sun' || val=='Sat'){
    return ('Weekend')
  }
  return ('Weekday')
}
fullData$weekday = as.factor( sapply(fullData$date,isWeekend))

aggData <-aggregate(as.matrix(fullData[,"steps"]), 
    as.list(fullData[,c("weekday","interval")]), 
    FUN=mean)

```

Now we aggregated data over interval and weekday

Visualizing differences in patterns between weekdays and weekends
```{r}
library("lattice")
xyplot(V1 ~ interval | weekday, data=aggData, type="l", layout=c(1,2))
```
We can see that peak usage happens at roughly the same time, though activity after peak usage time is different.

